### 2017-11-01 preamble.tex
add institution contribution table

\usepackage{authblk}
\usepackage{array}

\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}

\author[1]{Fran\c{c}ois Michonneau\thanks{francois.michonneau@gmail.com}}
\author[2]{Gustav Paulay\thanks{paulay@flmnh.ufl.edu}}
\author[1]{Mark Q. Martindale\thanks{mqmartin@whitney.ufl.edu}}

\affil[2]{Florida Museum of Natural History, University of Florida, Gainesville,
  FL 32611-7800, USA}

\affil[1]{Whitney Laboratory for Marine Bioscience, University of Florida, St Augustine, FL 32080, USA}

\renewcommand\Authands{ and }
\date{}

%%% Local Variables:
%%% mode: markdown
%%% TeX-master: "manuscript.tex"
%%% End: